---
title: "Predictive Model for Movie Revenue Estimation and Decision Support"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  #pdf_document:
    #toc: yes
    #toc_depth: '3'
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

Detailed Problem Statement:
The movie industry is a dynamic and complex environment where various factors influence a movie's success, particularly its revenue. This project focuses on developing a robust "Predictive Model for Movie Revenue Estimation and Decision Support." In an era of high uncertainty and financial risk in the film production business, understanding the determinants of movie revenue is paramount for informed decision-making, investment strategies, and marketing planning.

Movies are a unique combination of artistic expression and commercial ventures. To that end, this research investigates a multitude of features that contribute to the financial performance of a movie. These include aspects like genres, original languages, overview, popularity, production companies, release date, budget, runtime, status, tagline, vote average, vote count, credits, keywords, poster path, backdrop path, and recommendations. Each of these factors is scrutinized meticulously to unravel their individual and collective influence on movie revenue.

By conducting a comprehensive analysis of these features, this research aims to build a regression model (Linear, Decision Tree, Support Vector Machine) based on valuable insights through identifying patterns, relationships, and correlations. The findings of this study are expected to have substantial implications for movie production companies, investors, and stakeholders. They can leverage the predictive model developed in this project to make informed decisions regarding their movie investments, marketing strategies, and financial planning, ultimately contributing to the success and profitability of their ventures in the film industry.

**Importing the Movies Data**

```{r}
movies = read.csv("C:\\Anand\\Projects_GWU\\DATS6101-P2-Movie-Revenue-Prediction\\movies.csv\\Movie_data.csv")
#head(movies)
summary(movies)
movie = movies # Creating a copy of data frame
```

**Removing Unwanted Columns**

```{r}
library(dplyr)
remove_col <- c(1,5,7,13,16,17,18,19,20)
movie <- movie[, -remove_col]
head(movie)
```

**Summary of the data**

```{r}
new_column_names <- c("Title", "Genre", "Language", "Popularity", "Release_Date","Budget","Revenue","Runtime","Status","Avg_Vote","Vote_Count","Trailer_Views","Trailer_Likes")
colnames(movie) <- new_column_names
summary(movie)
```

Inferences:
1. Budget of Movies are having value 0, which is not possible (Remove rows with budget 0)
2. Revenue has negative values which is again not possible (Remove rows with revenue 0 or negative)
3. Runtime of Movies are having value 0, which is not possible (Remove rows with runtime 0)
4. Check for outliers and influential points in all the columns.

**Checking for missing values in the data**

```{r}
library(ggplot2)

# Calculate the percentage of missing values in each column
column_na_percentages <- colSums(is.na(movie)) / nrow(movie) * 100

# Set a smaller font size
smaller_font_size <- 0.8

# Create a bar plot for the percentage of missing values in each column
barplot(column_na_percentages, main = "Percentage of Missing Values in Each Column",
        xlab = "Columns", ylab = "Percentage of Missing Values",
        col = "lightcoral", border = "black", ylim = c(0, 100),
        names.arg = names(column_na_percentages), las = 2, cex.names = smaller_font_size)

print("Rows in Dataset before dropping rows with NA Values")
print(nrow(movie))

movie <- movie[complete.cases(movie), ]

print("Rows in Dataset After dropping rows with NA Values")
print(nrow(movie))
```

Removed rows with NA Values

***Data Cleaning***

```{r}
print("Rows in Dataset before cleaning")
print(nrow(movie))

# 1. Remove rows with Budget less than or equal to 0
movie <- movie[movie$Budget > 0, ]

# 2. Remove rows with Revenue less than or equal to 0
movie <- movie[movie$Revenue > 0, ]

# 3. Remove rows with Runtime less than or equal to 0
movie <- movie[movie$Runtime > 0, ]

print("Rows in Dataset After cleaning")
print(nrow(movie))
```

```{r}
selected_columns <- c("Popularity", "Budget", "Revenue", "Runtime", "Avg_Vote", "Vote_Count", "Trailer_Views","Trailer_Likes")

box_plots_list <- list()

# Create box plots for each selected column
for (col in selected_columns) {
  box_plot <- ggplot(movie, aes(y = get(col))) +
    geom_boxplot(fill = "lightblue", color = "black", alpha = 0.7) +
    labs(title = paste("Box Plot for", col), y = col) +
    theme_minimal() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.title = element_text(size = 12)) +
    theme(axis.text = element_text(size = 10))
  
  box_plots_list[[col]] <- box_plot
}

for (col in selected_columns) {
  print(box_plots_list[[col]])
}
```
**Data Visulisation**
[Scatter Plots, Correlation Heatmap etc]
```{r}
# Scatterplots
# Define the predictors and response variable
predictors <- c("Trailer_Likes", "Trailer_Views", "Vote_Count", "Budget", "Runtime", "Avg_Vote")

scatter_plots <- lapply(predictors, function(var) {
  ggplot(movie, aes(x = get(var), y = movie$Revenue)) +
    geom_point() +
    labs(title = paste("Scatter plot of", var, "vs. Revenue"))
})

# Display scatter plots
scatter_plots

```


**Stepwise Forward Feature Selection**
```{r}
library(MASS)

# Define your initial model
initial_model <- lm(Revenue ~ 1, data = movie)  # Simple model with intercept only

# Perform stepwise forward selection
final_model <- stepAIC(initial_model, direction = "forward", scope = list(lower = ~1, upper = ~Popularity + Budget + Runtime + Avg_Vote + Vote_Count + Trailer_Views + Trailer_Likes), data = movie)

# Display the final model summary
summary(final_model)

```

The features given in the final model using forward feature selection are Trailer_Likes, Trailer_Views, Vote_Count, Budget, Runtime ,Avg_Vote , popularity but we can see that popularity is not significant hence we will drop popularity and build the model again.

```{r}
lm_model_usingFFS <- lm(Revenue ~ Trailer_Likes + Trailer_Views + Vote_Count + Budget + Runtime + Avg_Vote, data = movie)

# Summarize the model
summary(lm_model_usingFFS)
```

In the above model we can see that all the input features are significant and we have a good Multiple R-Squared value. 
The p-value for the model is also significant indicating the model is a significant model.
The Residual Std Error is around 5.4 Million which seems a little high as we cannot exactly predict the Revenue of a Movie as it depends on various other factors.
```{r}
library(caret)

# Set the seed for reproducibility
set.seed(123)

# Define the predictors and response variable
predictors <- c("Trailer_Likes", "Trailer_Views", "Vote_Count", "Budget", "Runtime", "Avg_Vote")
response <- "Revenue"

# Create a training and testing dataset
train_index <- createDataPartition(movie$Revenue, p = 0.8, list = FALSE)
train_data <- movie[train_index, ]
test_data <- movie[-train_index, ]

# Fit the model on the training data
model <- lm(paste(response, "~", paste(predictors, collapse = " + ")), data = train_data)

# Make predictions on the test data
predictions <- predict(model, newdata = test_data)

# Calculate RMSE
rmse <- sqrt(mean((test_data$Revenue - predictions)^2))

# Print the RMSE
cat("Root Mean Squared Error (RMSE):", rmse, "\n")

```

```{r}
lm_model <- lm(Revenue ~ Trailer_Views+Trailer_Likes, data = movie)

# Summarize the model
summary(lm_model)
```

```{r}
library(caTools)

# Set a random seed for reproducibility
set.seed(123)

# Split the data into training (80%) and testing (20%) sets
split <- sample.split(movie$Revenue, SplitRatio = 0.8)
train_data <- subset(movie, split == TRUE)
test_data <- subset(movie, split == FALSE)

# Fit linear regression model on the training data
lm_model <- lm(Revenue ~ Trailer_Views+Trailer_Likes, data = train_data)

# Make predictions on the test data
predictions <- predict(lm_model, newdata = test_data)

# Calculate Root Mean Squared Error (RMSE)
rmse <- sqrt(mean((test_data$Revenue - predictions)^2))
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
```

```{r}
mean_revenue <- mean(movie$Revenue, na.rm = TRUE)
median_revenue <- median(movie$Revenue, na.rm = TRUE)
quantiles_revenue <- quantile(movie$Revenue, c(0.25, 0.5, 0.75), na.rm = TRUE)

# Print the results
cat("Mean Revenue:", mean_revenue, "\n")
cat("Median Revenue:", median_revenue, "\n")
cat("Quantiles (25%, 50%, 75%):", quantiles_revenue, "\n")
```

```{r}
```